Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-16 resex.tex
Revised text of random excitation results. Added note in README to verify magnitude of loss rates.

The changes in emittance are also directly visible in the beam
profiles measured with the BSRT. As an example
Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation
only in V in comparison to a reference bunch. The distribution also
stays Gaussian in case of a random excitation in contrast to the
\tenthtp\ in which case it clearly assumes a non-Gaussian shape (see
Fig.~\ref{fig:10thexpprof}).

The experimental results for a maximum excitation amplitude of
$A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm
the in general small losses predicted in simulations. The losses then
increase more dramatically around the time when the maximum excitation
amplitude is increased to $A_{\mathrm{max}}=12$~nrad (see
Fig.~\ref{fig:ranexpprof}). This is most likely the point when, due to
the large emittance growth, the physical aperture starts cutting into
the more central region of the distribution. Furthermore, the
dependence of the loss rate on the excitation plane is additive,
meaning that double the loss rate is observed for an excitation in H+V
compared to only in H or only V. The loss rate is also considerably
reduced for the bunches with transverse damper active independent of
the plane the excitation is applied in (see
Fig.~\ref{fig:ranexplossplane}). In comparison to the \seventhtp, the
loss rate for a random excitation is smaller for a maximum excitation
amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for
$A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}). This
indicates an increase of the diffusion mostly in the beam tails for
\seventhtp, while for a random excitation mostly the core is affected.

>>>>

The changes in beam distributions were also directly detected in the
BSRT profiles. As an example, Fig.~\ref{fig:ranexpprof} shows the
profiles induced by a vertical random excitation, compared to those of
a reference bunch. In this case, the distribution widens, but it
retains a Gaussian form, in contrast to the case of \tenthtp, when it
assumed a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}, for
instance).

The measured loss rates are presented in
Fig.~\ref{fig:ranexplossplane} as a function of random excitation
amplitude in H, V, and H+V planes.  Losses increased quadratically
with excitation amplitude, contributing approximately 3\%/h at
12~nrad. In simulations, the qualitative behavior was the same;
however, the predicted magnitude was much lower (0.1\%/h at
12~nrad). Moreover, the measured effect of the combined H+V excitation
was approximately the sum of H and V separately.

The systematic effects on losses during random excitations were
smaller than during resonant pulsing, as shown by the agreement
between the 3 data sets (black, red, yellow) within each of the
excitation modes in Fig.~\ref{fig:ranexplossplane}. This fact may be
an indication of the random excitation affecting the whole beam,
whereas, in the cases under study, resonant pulsing patterns excited
specific amplitude regions and were therefore more susceptible to
their population and to the order in which experiments were
performed. Another indication comes from the direct comparison of loss
rates vs.\ excitation amplitude for random and \seventhtp\
(Fig.~\ref{fig:ranexploss}). The first time it is applied (red
points), the 7th-turn excitation grows quadratically with amplitude,
whereas the second time it shows signs of saturation (yellow). In any
case, although the dynamics of their action may be different, these
two pulsing patterns were the most powerful, generating losses at the
level of several percent per hour at an amplitude of 12~nrad.

Another feature of random excitation was that the loss rates were
reduced by about a factor~2 when the transverse damping system was
active (dashed vs.\ solid lines in
Fig.~\ref{fig:ranexplossplane}). This was true for all three modes of
excitation (H, V, and H+V). On the other hand, for the resonant
pulsing modes, the transverse damper seemed to have a negligible
effect. This observation is discussed in more detail in
Section~\ref{sec:damp}.
